
\chapter*{Acknowledgments}

%\addcontentsline{toc}{chapter}{Acknowledgements} 

This doctoral thesis is the result of my research efforts over the last three years.  However, this work would not be possible without the collaboration and support of a wide number of colleagues, friends and institutions. First of all, I am infinitely grateful to my supervisors, Dr. Arnau Oliver and Dr. Xavier Lladó, for giving me the opportunity to work in this project. I really appreciate their enthusiasm, insight, unconditional support and friendship specially when difficulties arose. I owe them a lot for continuously encourage me not to stop forward. Without Arnau and Xavi, this doctoral thesis would simply not be possible.   

Several institutions and medical centers have been involved also.  I want to express my gratitude to the Generalitat de Catalunya for awarding me with the research grant FI-DGR2013, which has been used to fund this doctoral thesis. Most of the images used in this work have been gently facilitated by different research hospital centers in Catalunya. Furthermore, I want to thank to Dr. Lluís Ramió-Torrentà, Dr. Hector Perkal,  René Robles and Dr. Brigitte Beltrán from the Hospital Dr. Josep Trueta of Girona, Dr. Joan Carles Vilanova from the Hospital Santa Caterina of Girona, Dr. Deborah Pareto and Àlex Rovira from the Vall d’Hebrón research hospital of Barcelona and Dr. Jaume Sastre-Garriga from the Multiple Sclerosis Center of Catalunya for their support, continuous reviewing processes and patience answering our medical questions.  Moreover, I am sincerely grateful to all the reviewers that have been involved in the non-profited effort of reviewing not only this final manuscript but also each one of the research papers that compose it.

I am indebted with so many colleagues and friends for inspiring me with their work.  I would like to thank all the permanent staff, post-doctoral researchers, PhD students and administrative staff of the VICOROB team for their help during these years.  I am specially indebted with Aina, for her continuous efforts and patience helping me with my master’s grant stuff and documentation.  I want to express my appreciation to Dr. Jordi Freixenet and Dr. Joan Martí for their expert advices and the discussions about the medical image analysis field. I also want to thank the rest of my office colleagues Josep, Konstantin, Ricard P., Ricard C., Mojdeh, Eloi G., Ferran, Habib, Marc, Albert P., and Christian for their revealed patience and friendship.  Likewise, I would like to express my gratitude as well for Guillaume, Sik, Robert, Xevi and Yago for their technical discussions thorough the time we were running across the university's nearby forests. And very especially, I am very grateful to my colleagues Sandra, Mostafa, Onur, Mariano, and Eloy of the NeuroImaging Computing group for the inspiring discussions, for sharing their knowledge and helping me so much during these last few years.    

Not without a high probability to forget anyone, I want to express also my gratitude to my old friends Albert, Eloi, Miquel, Presser, Aida, Mònica, Clara, Sergio, Dani MC, etc… for being so close during these years.  Not to mention my old friends at FUNIAL: Moisès, Alfred, Antonio, Moi, Antonio C., Francesc, Dani, Edu, Mena, etc… for helping me to grow as a person.  Besides that, I will be forever indebted to Alicia, for her constant support, joy, and love during most of these years.   

However, this thesis would also not be possible without the unconditional love of my parents and my brother. Eight years ago, barely nobody but them were crazy enough to understand and support my decision of quitting my job position to start my degree in computer engineering. During these years, I have forgotten the number of times that their love and faith in me have been my light in the darkness. This doctoral thesis is entirely dedicated to them, whose sacrifice has always nourished my dreams.

Last but not least, I wish to thank to Mireia for her advice, support, daily infinite love, patience during the last months of intense work and for this smile that reminds me that this world is a very nice place to stay, when shared with her: like Marina and Núria, our coming daughter Helena will be very proud of you.  

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End:
