
\chapter*{Publications}

\addcontentsline{toc}{chapter}{Publications} 

The presented thesis is a compendium of the following research articles:

\begin{itemize}

\item  \textbf{Sergi Valverde}, Eloy Roura, Arnau Oliver, Eloy Roura, Sandra Gonz\'{a}lez, Deborah Pareto, Joan C. Vilanova, Llu\'{i}s Rami\'{o}-Torrent\`{a}, \`{A}lex Rovira and Xavier Llad\'{o}.  Automated brain tissue segmentation of MR images in the presence of white matter lesions. \textit{Submitted to Medical Image analysis}. 2016. Elsevier.  Quality index:  [JCR CSAI IF 3.654, Q1 (10/121)].

\item \textbf{Sergi Valverde}, Arnau Oliver, Eloy Roura, Deborah Pareto, Joan Carles Vilanova, Llu\'{i}s Rami\'{o}-Torrent\`{a}, Jaume Sastre-Garriga, Xavier Montalban, \`{A}lex Rovira and Xavier Llad\'{o}. Quantifying brain tissue volume in mutliple sclerosis with automated lesion segmentation and filling SPM8 toolboxes. \textit{NeuroImage: Clinical}. Vol. 9, pp 640-647. 2016. Elsevier.  Quality index:  [JCR N IF 2.526, Q2(6/14)].

\item  \textbf{Sergi Valverde}, Arnau Oliver, Yago D\'{i}ez, Mariano Cabezas, Joan Carles Vilanova, Llu\'{i}s Rami\'{o}-Torrent\`{a}, \`{A}lex Rovira, and Xavier Llad\'o. Evaluating the effects of white matter multiple sclerosis lesions on the volume estimation of six brain tissue segmentation methods. \textit{American Journal of Neuroradiology}. Vol. 36(6), pp. 1109-1115. 2015. American Society of Neuroradiology.  Quality index: [JCR RNMMI IF 3.589, Q1(19/125)].

\item \textbf{Sergi Valverde}, Arnau Oliver, Mariano Cabezas, Eloy Roura and Xavier Llad\'{o}. Comparison of ten brain tissue segmentation methods using revisited IBSR annotations. \textit{Journal of Magnetic Resonance Imaging}. Vol. 41, Issue 1, pp. 93-101. January 2015. Wiley.  Quality index: [JCR RNMMI IF: 3.210 Q1(23/125)].

\item  \textbf{Sergi Valverde}, Arnau Oliver, and Xavier Llad\'o. A white matter lesion-filling approach to improve brain tissue volume measurements. \textit{NeuroImage: Clinical}. Vol. 6, pp 86-92. 2014. Elsevier. Quality index: [JCR N IF 2.526, Q2(6/14)].


\end{itemize}


The rest of publications and conferences related with this PhD thesis are the following:

\section*{Journals}

\begin{itemize}

\item Eloy Roura, Nicolae Sarbu, Arnau Oliver, \textbf{Sergi Valverde}, Sandra Gonz\'{a}lez, Ricard Cervera, Nuria Bargall\'{o} and Xavier Llad\'{o}. Automated detection of Lupus white matter lesions in MRI images. \textit{Submited to Frontiers in Human Neuroscience}. 2016. Frontiers.  Quality index: [JCR PS IF 3.626, Q1(13/76)].

\item Eloy Roura, Arnau Oliver, Mariano Cabezas, \textbf{Sergi Valverde}, Deborah Pareto, Joan Carles Vilanova, Llu\'{i}s Rami\'{o}-Torrent\`{a}, \`{A}lex Rovira and Xavier Llad\'{o}. A toolbox for multiple sclerosis lesion segmentation. \textit{Neuroradiology}. Vol. 57(10),  pp. 1031-1043. 2015. Springer.  Quality index: [JCR RNMMI IF 2.485, Q2(41/125)].

\item Mariano Cabezas, Arnau Oliver, \textbf{Sergi Valverde}, Brigitte Beltr\'{a}n, Joan Carles Vilanova, Llu\'{i}s Rami\'{o}-Torrent\`{a}, \`{A}lex Rovira and Xavier Llad\'{o}. BOOST: a supervised approach for multiple sclerosis lesion segmentation. \textit{Journal of Neuroscience Methods}. Vol. 237, pp 108-117. 2014. Elsevier.  Quality index: [JCR N IF 2.025, Q3(174/252)].

\item Yago D\'{i}ez, Arnau Oliver, Mariano Cabezas, \textbf{Sergi Valverde}, Robert Mart\'{i}, Joan Carles Vilanova, Llu\'{i}s Rami\'{o}-Torrent\`{a}, \`{A}lex Rovira and Xavier Llad\'{o}. Intensity based methods for brain MRI longitudinal registration. A study on multiple sclerosis patients. \textit{Neuroinformatics}. Vol 12(3), pp 365-379. 2014. Springer.  Quality index: [JCR CSTM IF 2.825, Q1(13/102)] 

\end{itemize}


\section*{Conferences}

\begin{itemize}
\item Eloy Roura, Arnau Oliver, Mariano Cabezas, \textbf{Sergi Valverde}, Deborah Pareto, Joan Carles Vilanova, Llu\'{i}s Rami\'{o}-Torrent\`{a}, \`{A}lex Rovira and Xavier Llad\'{o}. An SPM12 extension for multiple sclerosis lesion segmentation. \textit{SPIE Medical Imaging 2016}. February 2016, San Diego, USA.

\item \textbf{Sergi Valverde}, Arnau Oliver, Eloy Roura, Deborah Pareto, Joan Carles Vilanova, Llu\'{i}s Rami\'{o}-Torrent\`{a}, Jaume Sastre-Garriga, Xavier Montalban, \`{A}lex Rovira and Xavier Llad\'{o}. Evaluation of two automated lesion segmentation and filling pipelines for brain tissue segmentation of multiple sclerosis patients. \textit{ECTRIMS 2015. Multiple Sclerosis}. October 2015, Barcelona, Spain.  Quality index: [JCR CN IF:4.472 Q1(25/191)].

\item Eloy Roura, Arnau Oliver, Mariano Cabezas, \textbf{Sergi Valverde}, Deborah Pareto, Joan Carles Vilanova, Llu\'{i}s Rami\'{o}-Torrent\`{a}, \`{A}lex Rovira and Xavier Llad\'{o}. A toolbox for segmenting multiple sclerosis lesions using T1w and FLAIR images. \textit{ECTRIMS 2015. Multiple Sclerosis}. October 2015, Barcelona, Spain.  Quality index: [JCR CN IF:4.472 Q1(25/191)].

\item \textbf{Sergi Valverde}, Arnau Oliver, Deborah Pareto, Joan Carles Vilanova, \`{A}lex Rovira , Llu\'{i}s Rami\'{o}-Torrent\`{a} and Xavier Llad\'{o}.
SLF: a MS white matter lesion filling toolbox for the SPM software. \textit{ECTRIMS 2014. Multiple Sclerosis}. September 2014, Boston, USA.  Quality index: [JCR CN IF:4.822 Q1(22/192)].

\item Ester Quintana, Brigitte Beltr\'{a}n, \textbf{Sergi Valverde}, Ren\'{e} Robles-Cedeno, Hector Perkal, Xavier  Llad\'{o}, Jos\'{e} Manuel Fern\'{a}ndez-Real and Llu\'{i}s Rami\'{o}-Torrent\`{a}. Expression of miRNAs in multiple sclerosis cerebrospinal fluid and their relation to MR activity. \textit{ECTRIMS 2014. Multiple Sclerosis}. September 2014, Boston, USA.  Quality index: [JCR CN IF:4.822 Q1(22/192)].

\item  Ester Quintana, Brigitte Beltr\'{a}n, \textbf{Sergi Valverde}, Ren\'{e} Robles-Cedeno, Hector Perkal, Xavier  Llad\'{o}, Jos\'{e} Manuel Fern\'{a}ndez-Real and Llu\'{i}s Rami\'{o}-Torrent\`{a}. Relaci\'{o}n entre la expresi\'{o}n de mirnas en LCR y variables de RM. \textit{Neurolog\'{i}a}, vol 29, pp 66-67. 2014.  Quality index: [JCR CN IF:1.322 Q3(142/191)].

\item \textbf{Sergi Valverde}, Arnau Oliver, Mariano Cabezas, Yago D\'{i}ez, Jordi Freixenet, Xavier Llad\'{o}, Joan Carles Vilanova, \`{A}lex Rovira and Llu\'{i}s Rami\'{o}-Torrent\`{a}. A quantitative study of the effects of White Matter MS lesions on tissue segmentation methods. \textit{ECTRIMS 2013. Multiple Sclerosis}. October 2013, Copenhagen, Denmark.  Quality index: [JCR CN IF:4.472 Q1(25/191)].



\end{itemize}



%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End:
